Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-26 sigcse-17/paper.tex
sigcse: some revisions
sigcse: dos2unix paper
sigcse: import text on 410 from edupar, with some violence
sigcse: save some space
sigcse: write about landlside
sigcse: add some pretty pictures
sigcse: write future work and fix some todos, fix abstract claims

\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}

@@

@@

% TODO camready
\newcommand\carnegiemellon{{\em Double-Blind University}\xspace}
\newcommand\cmu{DBU\xspace}
\newcommand\pebbles{\cmu-OS\xspace}
\newcommand\landslide{Landslide\xspace}
\newcommand\fourten{CS999}
%\newcommand\carnegiemellon{{\em Carnegie Mellon University}\xspace}
%\newcommand\cmu{CMU\xspace}
%\newcommand\pebbles{Pebbles\xspace}
%\newcommand\landslide{Landslide\xspace}
%\newcommand\fourten{15-410}

@@

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present Landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using Landslide to find previously-overlooked bugs in their own code, and discuss future methods for making Landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students Landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode Landslide requires manual instrumentation, testing thread libraries is fully automatic.
% TODO: verify these, obv. ;)
showing that it allows them to find and fix previously-overlooked bugs.
Further, %even struggling students can take advantage of it,
Landslide's users achieve higher concurrency scores on average in subsequent projects,
% FIXME awkward phrasing
suggesting that debugging with a model checker produces better student retention of concurrency principles.

@@

% TODO, do dis
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

@@

\input{intro}

\section{The {\secit Body} of The Paper}
% TODO
TODO body

\section{Conclusions}
% TODO
TODO conclusion

>>>>

\usepackage{xspace}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{enumitem}
%\usepackage{colortbl}

@@

\newcommand{\sect}[1]{\S #1}

\setlist[itemize]{leftmargin=*}

@@

% TODO camready
\newcommand\carnegiemellon{{\em Double-Blind University}\xspace}
\newcommand\cmu{DBU\xspace}
\newcommand\pebbles{\cmu-OS\xspace}
\newcommand\landslide{Landslide\xspace}
\newcommand\fourten{CS999\xspace}
\newcommand\twothirteen{CS888\xspace}
%\newcommand\carnegiemellon{{\em Carnegie Mellon University}\xspace}
%\newcommand\cmu{CMU\xspace}
%\newcommand\pebbles{Pebbles\xspace}
%\newcommand\landslide{Landslide\xspace}
%\newcommand\fourten{15-410}

@@

We aspire to expand the scope of these projects in response to
typical hardware platforms'
increasing complexity,
especially massive multi-core.
To that end, we wish to reduce the time students spend
debugging concurrency bugs.
We present Landslide, a
%systematic testing
stateless model checker which
offers a more principled approach than stress testing.
% edupar version
%We report on our preliminary experience working with students using Landslide to find previously-overlooked bugs in their own code, and discuss future methods for making Landslide more accessible to struggling students.
% sigcse version
We report on our experience giving students Landslide during the thread library
%and kernel projects.
project,
%Although kernel-mode Landslide requires manual instrumentation, testing thread libraries is fully automatic.
showing that it allows them to find and fix previously-overlooked bugs.
%% no :( :(
% Further, %even struggling students can take advantage of it,
% Landslide's users achieve higher concurrency scores on average in subsequent projects,
% suggesting that debugging with a model checker improves student retention of concurrency principles.

@@

% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
\begin{CCSXML}
	 <ccs2012>
	 <concept>
	 <concept_id>10011007.10011006.10011073</concept_id>
	 <concept_desc>Software and its engineering~Software maintenance tools</concept_desc>
	 <concept_significance>500</concept_significance>
	 </concept>
	 <concept>
	 <concept_id>10010147.10011777.10011778</concept_id>
	 <concept_desc>Computing methodologies~Concurrent algorithms</concept_desc>
	 <concept_significance>300</concept_significance>
	 </concept>
	 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software maintenance tools}
\ccsdesc[300]{Computing methodologies~Concurrent algorithms}

@@

\input{intro}
\input{410}
\input{landslide}
\input{eval}

%\section{Conclusions and Future Work}
\section{Future Work}

We are presently extending \landslide to support Pintos kernels \cite{pintos},
%used at many top CS schools including Stanford, Berkeley, and University of Chica
another instructional kernel popular at many top CS schools.
%a more widely-used instructional kernel than \pebbles.
Because Pintos mandates more internal APIs than \pebbles,
\landslide can more easily instrument student code without requiring manual annotations.
In a preliminary experiment, we tested 79 student Pintos implementations with \landslide,
finding 59 concurrency bugs with 10 CPU-hours allocated to each of 3 test cases \cite{quicksand-anonymized}.
In future work, we hope to fully automate Pintos instrumentation, and offer \landslide at other schools beyond \cmu.

%\fourten, the undergraduate operating systems course at \cmu, teaches concurrency in a
